mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2022-12-09 Draft3.Rmd
Draft 3

<<<<<<< HEAD
---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-11-09"
abstract: >
  This R package translates R objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic description, this new feature
  is a futher step towards reproducible science. The package can be extended by
  user-defined hooks that allow to replace variable names and new R functions by
  a suitable symbolic representation. We provide two working examples that
  illustrate the use of the new package in a scientific article and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/gondan-rochon/index.html.en
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/alfarone/index.html.en
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)

@@

The R extension of the markdown language [@Xie2020] enables reproducible
statistical reports with nice typesetting in HTML, Microsoft Word, and Latex.
However, so far, R expressions such as `pbinom(k, N, p)` are printed as-is; 
prettier mathematical expressions such as $P_\mathrm{Bi}(X \le k; N, p)$ require
Latex commands like `P_\mathrm{Bi}\left(X \le k; N, p\right)`. Except for
minimalistic examples, these commands are tedious to type in and hard to read in
the original code. Since recently [@R, version 4.2], manual pages include
support for mathematical expressions [@Sarkar2022], which already is a big
improvement. However, rules for the translation of R's built-in language elements to
their mathematical representation are still lacking.

In this R package, such rules are defined for the _automatic_ translation of R
expressions mathematical output. The translation is done by an embedded
Prolog interpreter that maps nested expressions recursively to MathML and
LaTeX/MathJax, respectively, for typesetting of R Markdown documents in
different formats or ShinyApp webpages. User-defined hooks enable to extend the
set of rules and/or replace specific R elements by special mathematical signs.

The paper is organized as follow. Section 2 presents the background of the mathml package, namely Prolog and the R package \pkg{rolog}. Section 3 illustrates the features of the \pkg{mathml} package, providing some examples. Section 4 embeds the Discussion and the presentation of a Case Study. Section 5 is devoted to Conclusions and Further Research. The reader may also find some additional characteristics of the package and its implementation in the Appendix.

@@

@@

=======
---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-12-09"
abstract: >
  This R\ package translates R\ objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R\ functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R\ code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic representation, this new
  feature is a futher step towards reproducible science. User-defined hooks
  allow to extend the package by mapping new variable names and functions to
  new MathML and LaTeX entities. We provide a few working examples that
  illustrate the use of the package in a little scientific report and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/gondan-rochon/index.html.en
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/alfarone/index.html.en
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r, results="asis"}
t <- quote(omit_left(a + b))
mathout(t)

catmath(omit_left(a + b))
```

```{r, results="asis"}
t <- quote(omit_right(a + b))
mathout(t)

catmath(omit_right(a + b))
```

```{r, results="asis"}
t <- quote(add_left(a + b))
mathout(t)

catmath(add_left(a + b))
```

```{r, results="asis"}
t <- quote(add_right(a + b))
mathout(t)

catmath(add_right(a + b))
```

```{r, results="asis"}
t <- quote(add_right(a + b))
mathout(t)

catmath(add_right(a + b))
```

```{r, results="asis"}
t <- quote(instead(a, b) + c)
mathout(t)

catmath(instead(a, b) + c)
```

>>>>>>> 9ed4866198fb10109aebbba2714ff06f1e6e9df5

>>>>

---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-12-09"
abstract: >
  This R\ package translates R\ objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R\ functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R\ code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic representation, this new
  feature is a futher step towards reproducible science. User-defined hooks
  allow to extend the package by mapping new variable names and functions to
  new MathML and LaTeX entities. We provide a few working examples that
  illustrate the use of the package in a little scientific report and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/gondan-rochon/index.html.en
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.uibk.ac.at/psychologie/mitarbeiter/alfarone/index.html.en
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

@@

@@

```{r, results="asis"}
# It prints the incorrect version
mathout(t, flags=quote(error(asis)))
```

@@
